---
title: "MMA860_RDOTA2_Repo"
author: "Lin Ma"
date: "2022-03-22"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  cache=FALSE, 
  fig.width = 5, 
  fig.height = 4
  )
```


```{r libraries}
library(RDota2)
library(data.table)

library(tidyverse)
library(readr)
library(stringr)
library(reshape2)
library(readxl)
library(pander)
library(magrittr)
library(forcats)
library(data.table)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(pheatmap)
library(ggbeeswarm)

library(ape)

library(data.table)
library(ComplexHeatmap)
library(RColorBrewer)

library(ggsci)

library(viridis)
library(lubridate)

options(scipen=999)
options(digits=12)
panderOptions('knitr.auto.asis', FALSE)
panderOptions('knitr.auto.asis', FALSE)
panderOptions('digits', 6)
panderOptions('round', 6)


library(ggrepel)

options(scipen=999)
panderOptions('knitr.auto.asis', FALSE)

specify_decimal <- function(x, k) as.numeric(trimws(format(round(x, k), nsmall=k)))
```


# Introduction

- Data driven analysis is not make up something and try to let the data fit in. 
- Overall terminology is to look at  descriptive figures and statistics of our data set and try to make assumptions based on the game regulations and competition experiences. 

# Data Pre-processing

The first step was done in another file, refer to the github link [1]

## Load Data from Match ID: 3097027819

We pre-parse all the matches with ID: 3097027819 ... ( what kind of data - two major teams with 114 heros in total; Game type; **Killing scores and it relevant assumption(We found that the killing score is not representing the final winning team); winner team (e.g Rwin - not necessary mean that the killing score is higher than the team; etc)

```{r, echo=TRUE}
mid = "3097027819"
raw_game_data <- readRDS(paste0("RDS", mid)) %>%
  dplyr::rename(game_id = mid)
print(nrow(raw_game_data))
```

## Data Filtering 

Add explanation here: by looking at the data, we observed ... we need to figure out the most played game type. since we have different strategy per game type within different hero selections. since Different Confounding factor involved. 

```{r} 
# only keep match with more than 900 seconds
raw_game_data %<>% filter(duration >= 900) %>% unique()

# Extract radiant info: to simplicity, ignore the player order. (from raw to Radiant wide with game information, including Rscore, Dscore, Rwin.etc )
radiant_wide_df <- raw_game_data %>%
  select(game_id, R1:R5) %>% 
  gather(player_team, hero_id, R1:R5) %>% # wide to long 
  mutate(player_team = 1) %>% 
  spread(hero_id, player_team, fill = 0)  # long to wide
colnames(radiant_wide_df) <- c("game_id", paste("R", setdiff(colnames(radiant_wide_df), "game_id"), sep=""))

# Extract dire info
dire_wide_df <-raw_game_data %>% 
  select(game_id, D1:D5) %>% 
  gather(player_team, hero_id, D1:D5) %>% # wide to long
  mutate(player_team = 1) %>% 
  spread(hero_id, player_team, fill = 0)  # long to wide
colnames(dire_wide_df) <- c("game_id", paste("D", setdiff(colnames(dire_wide_df), "game_id"), sep=""))

# Filter match data based on game type 
raw_game_data %<>%
  mutate(game_type = ifelse(lobby == 1 & gmode == 2, "PRO", NA)) %>%
  mutate(game_type = ifelse(lobby == 7 & gmode == 2, "RCM", as.character(game_type))) %>%
  mutate(game_type = ifelse(lobby == 7 & gmode == 22, "RAP", as.character(game_type))) %>%
  mutate(game_type = ifelse(lobby == 7 & gmode == 3, "RRD", as.character(game_type)))
raw_game_data %>%
  dplyr::count(game_type) %>%
  pander(caption = "match counts for different game types")

# We only consider RAP game type
raw_game_data %<>% filter(game_type == "RAP") %>% select(-game_type)
print("the number of games left")
print(nrow(raw_game_data))


# Combine all the information via the unique ID: game_id
game_data_df <- raw_game_data %>% 
  select(game_id, duration, Rwin, Rscore, Dscore) %>%
  mutate(Rwin = as.integer(Rwin)) %>%
  left_join(radiant_wide_df, by=c("game_id")) %>% 
  left_join(dire_wide_df, by=c("game_id")) %>%
  mutate(Rwin_label = ifelse(Rwin == 1, "Radiat Win", "Dire Win"))
```


- **game_data_df**

## The relationship between Rscore, Dscore, and Rwin

GOAL: "we could convert our target variables into continuous value"

To simplify our problem, we only consider the winning games `Rwin == 1` where `Rscore` > 'Dscore`, and there are 750 games not meeting our assumption, therefore not included in the following analysis.

```{r}
game_data_df

game_data_df %>%
  ggplot(aes(x = Dscore, y = Rscore, color = Rwin_label)) + 
  geom_point(size = 1) + 
  scale_color_lancet() + 
  geom_abline(intercept = 0, slope = 1) + 
  theme_bw()

game_not_meet_assumption <- game_data_df %>% 
  mutate(delta = ifelse(Rscore > Dscore, 1, 0)) %>%
  filter(delta != Rwin)
print(nrow(game_not_meet_assumption))


game_data_df %<>% filter(!game_id %in% game_not_meet_assumption$game_id)
print(nrow(game_data_df))
```


# Descriptive Statistics About Game Data Dataframe

## Relationship between game duration and Rscore

```{r}
game_data_df %>% 
  select(game_id, duration, Rscore) %>%
  ggplot(aes(x = Rscore, y = duration)) + 
  geom_point() +
  theme_bw() + 
  labs(x = "", y = "Game duration")
```


## Feature selection: select player basd on the accumulative winning game counts

There are 114 heroes, each team select 5 heros at the beginning of the game. We start from looking at the attendance frequency of the heroes from both team. (Explanation needed) 

```{r}
attendence_df <- game_data_df %>%
  filter(Rwin == 1) %>% #<--- only look at the game where Radiant win
  gather(hero, attend, R1:R114) %>% group_by(hero) %>% summarise(total_winning_times = sum(attend)) 

attendence_df %>%
  ggplot(aes(x = total_winning_times)) + 
  geom_histogram()+
  geom_vline(xintercept = 400, color = "red") +
  theme_bw() +
  ggtitle("Hero attendence with Radiant win") + 
  theme(plot.title = element_text(hjust = 0.5))

nheros <- nrow(attendence_df %>% filter(total_winning_times >= 400))
top_radiants <- attendence_df %>% filter(total_winning_times >= 400) %>% .$hero
attendence_df %>% filter(total_winning_times >= 400) %>% 
  arrange(desc(total_winning_times)) %>% 
  pander(caption = paste("There are", nheros, "heros attend more than 400 games in the current match when Radiant win"))
```

We also looked at the most frequent attendance of heroes with dire team win.

```{r}
attendence_df <- game_data_df %>% 
  filter(Rwin == 0) %>%
  gather(hero, attend, D1:D114) %>% group_by(hero) %>% summarise(total_winning_times = sum(attend)) 

attendence_df %>%
  ggplot(aes(x = total_winning_times)) + 
  geom_histogram()+
  geom_vline(xintercept = 400, color = "red") +
  theme_bw() + 
  ggtitle("Hero attendence with Dire win") + 
  theme(plot.title = element_text(hjust = 0.5))
 
nheros <- nrow(attendence_df %>% filter(total_winning_times >= 400))
top_dires <- attendence_df %>% filter(total_winning_times >= 400) %>% .$hero
attendence_df %>% filter(total_winning_times >= 400) %>% 
  arrange(desc(total_winning_times)) %>% 
  pander(caption = paste("There are", nheros, "heros attend more than 400 games in the current match when Radiant win"))
```

** conclusion ** 

## Heatmap of all the heros with more than 400 winning games.

```{r, fig.width=14, fig.height=8}
top_gamers_df <- game_data_df %>% select(game_id, all_of(top_radiants), all_of(top_dires), Rwin_label) %>% 
  arrange(Rwin_label)
top_gamers_df

top_gamers_mat <- top_gamers_df %>% select(-one_of(c("game_id", "Rwin_label"))) %>% as.matrix()
rownames(top_gamers_mat) <- top_gamers_df$game_id

game_annotation <- top_gamers_df %>% select(Rwin_label)
rownames(game_annotation) <- top_gamers_df$game_id

pheatmap(top_gamers_mat, color = c("white", "red"), show_rownames = F, breaks = c(0, 0.5, 1), annotation_row = game_annotation)
```

- rows and columns are clustered via hierarchical cluster. 
And we don't observe any trend between the top 14 most frequent game winners and whether their teams win. 

- note: from the heat map, we don't see a clear trend between Rwin/Rlose with the top games. Besides, we need to filter games with no top gamer attending, otherwise conflicting information.

## Top gamers only 

Filter out games with no top gamer attending.

- top_gamers_df

```{r}
top_gamers_df <- game_data_df %>% select(game_id, all_of(top_radiants), all_of(top_dires), Rwin_label, Rscore, Dscore, duration) 
top_gamers_df

top_stats <- top_gamers_df %>% 
  select(game_id:R74) %>%
  gather(hero, attend, R104:R74) %>% 
  group_by(game_id) %>% 
  summarise(top_hero_counts = sum(attend)) %>% 
  ungroup() 
top_stats %>% ggplot(aes(x = top_hero_counts)) + 
  geom_histogram() + theme_bw() + 
  labs(x = "top hero attendence counts", y = "Radiant game counts")

top_gamers_df %<>% filter(game_id %in% unique(top_stats %>% filter(top_hero_counts > 0) %>% .$game_id))

print(nrow(top_gamers_df))
```


```{r}
top_stats <- top_gamers_df %>% 
  select(game_id, D104:D8) %>%
  gather(hero, attend, D104:D8) %>% 
  group_by(game_id) %>% 
  summarise(top_hero_counts = sum(attend)) %>% 
  ungroup() 
top_stats %>% ggplot(aes(x = top_hero_counts)) + 
  geom_histogram() + theme_bw() + 
  labs(x = "top hero attendence counts", y = "Dire game counts")

top_gamers_df %<>% filter(game_id %in% unique(top_stats %>% filter(top_hero_counts > 0) %>% .$game_id))

```

## Balanced data for the modeling

```{r}
top_gamers_df %>% 
  dplyr::count(Rwin_label) %>% 
  dplyr::rename(Game_counts = n) %>% 
  pander(caption = "balanced data for the model")

top_gamers_df %>% write.table("top_gamers_df.tsv", sep = "\t", quote = F, row.names = F)
```

- `top_gamers_df.tsv`

# Linear Regression Model

For the downstream analysis, we have 3010 games with balanced Rwin and Dwin. We also narrow down the games which meet the assumption that Rwin where `Rscore` > `Dscore`.

## Simpler model 

We start our modeling with only one target variable `duration` and predictor variable `Rscore`.

Rscore = a + duration * b

### Distribution of the match `duration`

```{r, fig.width = 5, fig.height = 4}
top_gamers_df %>%
  ggplot(aes(x = duration/60)) + 
  geom_histogram() +
  labs(x = "Match duration (minuts)", y = "Match counts") + 
  theme_bw() + 
  ggtitle("Distributuion of match duration for\nmatch id: 3097027819") + 
  theme(plot.title = element_text(hjust = 0.5))
```

### Virsulisation of `duration` with `Rscore`

```{r}
top_gamers_df %>%
  ggplot(aes( x= duration, y = Rscore)) + 
  geom_point() + 
  labs(x = "game duration (seconds)", y = "Rscore") + 
  theme_bw()
```

### Linear Regression Modeling 

Add explanation of LM result, refer to [2]

```{r, echo =T}
set.seed(101)
lm.model.simple <- lm(formula = Rscore ~ duration, data = top_gamers_df)

summary(lm.model.simple)
```

### Add linear regression back to the figure

```{r}
top_gamers_df %>%
  ggplot(aes( x= duration, y = Rscore)) + 
  geom_point() + 
  labs(x = "game duration (seconds)", y = "Rscore") + 
  theme_bw() + 
  geom_abline(slope = coef(lm.model.simple)["duration"], intercept = coef(lm.model.simple)["(Intercept)"], color = "red")
```


## Linear regression with selected radint player

### Heatmap of all the top most Radiant wining heros with more than 400 wining games.

- red: in the game
- white: not in the game

Because we only looked at the top winning R gamers, not all the games would have all top heroes participants. Therefore we take a look at the heatmap first. 

```{r, fig.width=14, fig.height=8}
my_df <- top_gamers_df %>% select(game_id, all_of(top_radiants), Rwin_label) %>% 
  arrange(Rwin_label)

my_mat <- my_df %>% select(-one_of(c("game_id", "Rwin_label"))) %>% as.matrix()
rownames(my_mat) <- my_df$game_id

game_annotation <- my_df %>% select(Rwin_label)
rownames(game_annotation) <- my_df$game_id

pheatmap(my_mat, color = c("white", "red"), show_rownames = F, breaks = c(0, 0.5, 1), annotation_row = game_annotation)
```

### Pairwise correlation coefficient between R top gamers

We can observe that the top R gamers participants are independent from each other [3]

```{r correlation coefficient}
library(corrplot)

my_df <- top_gamers_df %>% 
  select(R104:R8)
corrplot(cor(my_df), method = 'number')
```

### Linear Regression Model

Add explanation here. (Input - top heros selected and game duration; and their relationship with Rscore. game duration, and character of the heros- by name)

```{r}
set.seed(101)
lm.model.selected.player <- lm(formula = Rscore ~ duration + R104 + R14 + R2 + R35 + R44 + R74 + R8, data = top_gamers_df)

summary(lm.model.selected.player)
```


# References

[1] https://github.com/chengjunhou/Tutorial/blob/master/rdota2/dota2_data_query.R

[2] https://www.datacamp.com/community/tutorials/linear-regression-R

[3] http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

[4] http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

[5] https://stats.stackexchange.com/questions/25804/why-would-r-return-na-as-a-lm-coefficient

[6] https://stackoverflow.com/questions/7337761/linear-regression-na-estimate-just-for-last-coefficient